---
title: "A Simulation Study using Piecemeal"
author: Pavel N. Krivitsky
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Simulation Study using Piecemeal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Suppose that we want to simulate a function of two variables, $x$ and
$y$, for $x = 1, 2$ and $y = 1, 3, 9, 27$.

The function ultimately returns $x\times y$ and a random number
between 0 and 1, but to get there, it relies on an external constant
`a` equalling to 8 that is defined in its environment and on the
package `rlang`.

Unfortunately, our initial implementation also has some bugs that
cause it to crash if the sum of the product and the hundredths digit
of the random number is divisible by 4, with an error message that
depends on the remainder of dividing it by 8.

Here it is:
```{r}
a <- 8
f <- function(x, y) {
  p <- x*y
  u <- runif(1)

  errcond <- p + floor(u * 100) %% 10 + a
  if(errcond %% 4 == 0) stop("condition ", errcond %% 8, call. = FALSE)

  # rlang::dbl
  dbl(p = p, u = u)
}
```

We begin by setting the directory for run outputs. We will use R's temporary directory:
```{r}
outdir <- file.path(tempdir(), "piecemeal_demo")
```

Initialise the `Piecemeal` object.
```{r}
sim <- piecemeal::Piecemeal$new(outdir)
```

The following will clear the simulation outputs. Called without arguments, it will ask for confirmation first.
```{r}
sim$reset(FALSE)
```

The cluster has 2 nodes. We can also pass a cluster object:
```{r}
sim$cluster(2)
```
Set a factorial design with $x = 1, 2$ and $y = 1, 3, 9, 27$:
```{r}
sim$factorial(x = 2^(0:1), y = 3^(0:3))
```
Use 3 replications per combination of $x$ and $y$:
```{r}
sim$nrep(3)
```
Set the calling a function of x and y:
```{r}
sim$worker(f)
```

Note that we could have chained all these together:
```{r, eval = FALSE}
sim$
  cluster(2)$
  factorial(x = 2^(0:1), y = 3^(0:3))$
  nrep(3)$
  worker(f)
```

We can obtain a list of configurations (combinations of treatment and seed) to be run. (Here, first two are shown.)
```{r}
head(sim$todo(), 2)
```

We can now execute this setup. It will print a summary of the results of each run.
```{r}
sim$run()
```

Looks like all configurations failed, because the worker nodes
can't find the variable `a`. Let's export it to the worker nodes and try again:
```{r}
sim$export_vars("a")
sim$run()
```

We have new errors. Most of the nodes complain that we haven't
loaded the `rlang` package. Let's fix that:
```{r}
sim$setup({library(rlang)})
sim$run()
```

Some runs have succeeded. Here's what the individual run output files look like:
```{r}
list.files(outdir, recursive = TRUE)
```

Notice that they are split up into subdirectories. This improves
performance on some file systems. It can be controlled by the
`$options(split=)` setting.

If we run again, successful runs will be skipped:
```{r}
sim$run()
```

At any time, we can obtain the data frame of successful runs. If your
configuration or output data structure is more complex, you may need
to use custom functions for `result_df()`.
```{r}
sim$result_df()
```

The remaining errors are more subtle. We can examine the
configurations still to be done to see which did not succeed? (First
two are shown here.)
```{r}
head(sim$todo(), 2)
```

Alternatively, we can have the unsuccessful runs saved and inspect them:
```{r}
sim$options(error = "save")
sim$run()
head(sim$erred(), 2)
```

In fact, we can easily test-run the function to reproduce the error:
```{r, error = TRUE}
errcfg <- sim$erred()[[1]]
# debugonce(f) # to step through it
set.seed(errcfg$seed)
do.call(f, errcfg$treatment)
```

Let's say that based on the above, we were able to fix the bug. Our new functions is as follows:
```{r}
f <- function(x, y) {
  p <- x*y
  u <- runif(1)
  dbl(p = p, u = u)
}
```

Replace the worker function, clear the erred outputs, and run again.
```{r}
sim$worker(f)$clean()$run()
```

Now, we have all 24 combinations!
```{r}
sim$result_df()
```

Lastly, suppose that we want to run additional 2 replications of
each treatment combination. We can pick up where we left off:
```{r}
sim$nrep(5)
sim$run()
sim$result_df()
```
